Data Science Applied to Ag - Final Project - ML

### 2025-04-26 code/Draft_script.qmd
d

## 2. Random Forest Training  
### a. Random Forest Model specification

@@

rf_grid_result <- tune_sim_anneal(object = rf_spec,
                     preprocessor = weather_recipe,
                     resamples = resampling_foldcv,
                    #param_info = rf_param,
                    iter = 10 
                     )

@@

We want to minimize the RMSE. The lower the RMSE, the better. In the legend section of the RMSE grid plot seen above, the lowest RMSE is"2.85" which has a dark purple color, and we see the dark purple colored dot is on 4 th iteration, which indicates that the lowest RMSE [= "2.85"] was obtained in the 4 th iteration.

What tree_depth and mtry values created lowest RMSE?

tree_depth = ~ 110; mtry = ~ 22 [when we have 10 iterations]

@@

```{r R2}
rf_grid_result %>%
  collect_metrics() %>%
  filter(.metric == "rsq") %>%
  ggplot(aes(x = mtry, 
             y = trees 
             )) +
  geom_path(group = 1) +
  geom_point(aes(color = mean),
             size = 3) + 
  scale_color_viridis_b() +
  geom_text(aes(label = .iter), nudge_x = .0005) +
  labs(title = "R2")

@@

We want to see a maximized R2. The higher the R2, the better. In the legend section of the R2 grid plot seen above, the highest R2 is "0.16" which has a yellow color, and we see the yellow colored dot is on 4 th iteration, which indicates that the highest R2 [= "0.16"] was obtained in the 4 th iteration.

What tree_depth and min criterion values created highest R2?

tree_depth = ~110; mtry = ~ 22 [this was the best result for 10 iterations]

> Previously, we selected the single best model. Now, let's select the best model by "percent loss" of the metric, so we choose a model among the top ones that is more parsimonious.

```{r}
# Based on lowest RMSE
best_rmse <- rf_grid_result %>%
  select_by_pct_loss("trees",
                     metric = "rmse",
                     limit = 2 #"limit = 2": this is how many percent losses you are accepting, 2% in this case
                     )

@@

```{r}
# Based on greatest R2
best_r2 <- rf_grid_result %>%
  select_by_pct_loss("trees",
                     metric = "rsq",
                     limit = 2
                     )

>>>>

## 2. Random Forest Training

### a. Random Forest Model specification

@@

rf_grid_result <- tune_sim_anneal(object = rf_spec,
                     preprocessor = weather_recipe,
                     resamples = rf_resampling_foldcv,
                    #param_info = rf_param,
                    iter = 10 
                     )

@@

@@

```{r R2}

rf_grid_result %>%
  collect_metrics() %>%
  filter(.metric == "rsq") %>%
  ggplot(aes(x = mtry, 
             y = trees 
             )) +
  geom_path(group = 1) +
  geom_point(aes(color = mean),
             size = 3) + 
  scale_color_viridis_b() +
  geom_text(aes(label = .iter), nudge_x = .0005) +
  labs(title = "R2")

@@

The following code chunk will select the best model based on "percent loss" of the metric.

# Based on lowest RMSE
best_rmse <- rf_grid_result %>%
  select_by_pct_loss("trees",
                     metric = "rmse",
                     limit = 2 #"limit = 2": this is how many percent losses you are accepting, 2% in this case
                     )

@@

The following code chunk will select the best model based on greatest "R2"

# Based on greatest R2
best_r2 <- rf_grid_result %>%
  select_by_pct_loss("trees",
                     metric = "rsq",
                     limit = 2
                     )
